---
title: "Threshold for Ag Clone Calling"
vignette: >
  %\VignetteIndexEntry{Benchmarking}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
author: "`r 'Author: Matthew Hung'`"
date: "`r paste('Last Updated:', format(Sys.Date(), '%Y/%m/%d'))`"
output:
  html_document:
    code_folding: hide
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

This section is designed to test optimal minimum levenshtein distance cutoff to call "antigen specific" BCR clones. 

We have taken S2+ and S2- B cells from Lopez de Azziz et al. as positive and negative controls to determine the optimal cutoff in calling "antigen specific" BCR clones. To do this, we have ran `find_agBCR` on:

* TEST 1 : S2- memory B cells against S2+ memory B cells as reference
* TEST 2 : Total plasma cells against S2+ memory B cells as reference
* TEST 3 : Total plasma cells against S2- memory B cells as reference


```{r}
library(tidyverse)
library(ggprism)
library(ggridges)
```

Now we will load the output from `find_agBCR` for TEST 1. We will use levenshtein distance as method to calculate CDR3 distance.
```{r}
s2p_neg_vs_pos <- read.csv("./benchmark/test1/agBCR_by_heavyVJCDR3.csv") %>% 
    filter(heavyCDR3_dist_method == "levenshtein") %>%
    mutate(samples = "S2-_vs_S2+")
```

By plotting the distribution of minimum levenshtein distances, we can see only ~0.6% of the S2- negative cells (368/58714 cells) have a CDR3 distance that is less than 0.1.
```{r}
# set minimum levenshtein distance cutoff at 0.1
cutoff <- 0.1

# plot distribution of minimum levenshtein distances
options(repr.plot.width=5, repr.plot.height=4)
plot1 <- s2p_neg_vs_pos %>%
  ggplot(aes(x = heavyCDR3_dist, y = samples, fill = samples)) +
    ggridges::geom_density_ridges(alpha = 0.5) +
    geom_text(
      data = . %>%
        mutate(type = ifelse(heavyCDR3_dist < cutoff, "Binding", "Non-binding")) %>%
        group_by(type, samples) %>%
        summarize(count = n()) %>%
        group_by(samples) %>%
        mutate(pct = count*100/sum(count)) %>%
        mutate(label = paste0(type, " : ", round(pct, 1), " %\n (", count, "/", sum(count), ")")) %>%
        filter(type == "Binding"),
      aes(x = 0.1, y = samples, label = label),
      size = 4) +
    ggprism::theme_prism(border = T) + 
    scale_y_discrete(expand = c(0.5, 0.5)) +
    geom_vline(xintercept = cutoff, linetype = "dashed", color = "black") +
    labs(x = "heavy chain CDR3 distance", y = NULL, title = "Minimum levenshtein distance\nbetween S2- vs S2+ B cells")
print(plot1)
```

Now we will see if the same threshold (minimum levenshtein distance of 0.1) can be used to predict S2-specific clones in total plasma cells. Here we will load the output from `find_agBCR` for TEST 2 and TEST 3.
```{r}
PC_s2p_pos <- read.csv("./benchmark/test2/agBCR_by_heavyVJCDR3.csv") %>% 
    filter(heavyCDR3_dist_method == "levenshtein") %>%
    filter(str_detect(barcodes, "^Plasmablast_v1D14|^Plasmablast_v2D6")) %>%
    mutate(samples = "PC_vs_S2+")

PC_S2 <- bind_rows(list(s2p_neg_vs_pos, PC_s2p_pos)) %>%
  mutate(samples = factor(samples, levels = c("S2-_vs_S2+", "PC_vs_S2+")))
```

The distribution of minimum levenshtein distances show a bimodal distribution by comparing CDR3 distance of plasma cells with S2+ memory B cells, as compared to the unimodal distribution in comparison with S2- memory B cells. 

Setting a minimum levenshtein distance cutoff of 0.1 called 14.3% (1613/11254) S2+ plasma cells.
```{r}
options(repr.plot.width=5, repr.plot.height=4)
plot2 <- PC_S2 %>%
  ggplot(aes(x = heavyCDR3_dist, y = samples, fill = samples)) +
    ggridges::geom_density_ridges(alpha = 0.5) +
    geom_text(
      data = . %>%
        mutate(type = ifelse(heavyCDR3_dist < cutoff, "Binding", "Non-binding")) %>%
        group_by(type, samples) %>%
        summarize(count = n()) %>%
        group_by(samples) %>%
        mutate(pct = count*100/sum(count)) %>%
        mutate(label = paste0(type, " : ", round(pct, 1), " %\n (", count, "/", sum(count), ")")) %>%
        filter(type == "Binding"),
      aes(x = 0.1, y = samples, label = label),
      size = 4) +
    ggprism::theme_prism(border = T) + 
    scale_y_discrete(expand = c(0.5, 0.5)) +
    geom_vline(xintercept = cutoff, linetype = "dashed", color = "black") +
    labs(x = "heavy chain CDR3 distance", y = NULL, title = "Minimum levenshtein distance")
print(plot2)
ggsave("./benchmark/heavyVJ/Plasma_cell_vs_S2+.png", plot2, width = 8, height = 6)
```


Now we will see the distribution of minimum levenshtein distances of the S2-binding plasma cells with S2- memory B cells.
```{r}
PC_s2p_neg <- read.csv("./benchmark/test3/agBCR_by_heavyVJCDR3.csv") %>% 
    filter(heavyCDR3_dist_method == "levenshtein") %>%
    mutate(samples = "S2+ PC_vs_S2-") %>%
    filter(barcodes %in% (PC_s2p_pos %>% filter(heavyCDR3_dist < cutoff) %>% .$barcodes))

PC_S2 <- bind_rows(list(s2p_neg_vs_pos, PC_s2p_pos, PC_s2p_neg)) %>%
  mutate(samples = factor(samples, levels = c("S2-_vs_S2+", "PC_vs_S2+", "S2+ PC_vs_S2-")))
```

~10% (162/1613) of S2-binding plasma cells have similar CDR3 sequences to S2- memory B cells.
```{r}
options(repr.plot.width=5, repr.plot.height=4)
plot3 <- PC_S2 %>%
  ggplot(aes(x = heavyCDR3_dist, y = samples, fill = samples)) +
    ggridges::geom_density_ridges(alpha = 0.5) +
    geom_text(
      data = . %>%
        mutate(type = ifelse(heavyCDR3_dist < cutoff, "Binding", "Non-binding")) %>%
        group_by(type, samples) %>%
        summarize(count = n()) %>%
        group_by(samples) %>%
        mutate(pct = count*100/sum(count)) %>%
        mutate(label = paste0(type, " : ", round(pct, 1), " %\n (", count, "/", sum(count), ")")) %>%
        filter(type == "Binding"),
      aes(x = 0.1, y = samples, label = label),
      size = 4) +
    ggprism::theme_prism(border = T) + 
    scale_y_discrete(expand = c(0.5, 0.5)) +
    geom_vline(xintercept = cutoff, linetype = "dashed", color = "black") +
    labs(x = "heavy chain CDR3 distance", y = NULL, title = "Minimum levenshtein distance")
print(plot3)
ggsave("./benchmark/heavyVJ/Plasma_cell_vs_S2+-.png", plot3, width = 8, height = 6)
```


It is possible that these 10% of S2 binding plasma cells sharing similar CDR3 sequence with S2- cells are due to inadequate sorting of S2- cells.

```{r}
# get CDR3 sequences of S2- memory B cells that are similar to S2+ memory B cells
similar_cdr3 <- s2p_neg_vs_pos %>%
  filter(heavyCDR3_dist < cutoff) %>%
  distinct(heavyV, heavyJ, heavyCDR3, .keep_all = T) %>%
  .$heavyCDR3

# see how many of the 10% S2 binding plasma cells align to those S2- memory B cells
PC_s2p_neg %>%
    filter(heavyCDR3_dist < cutoff) %>%
    filter(ref_heavyCDR3 %in% similar_cdr3)
```

```{r}
reference <- get_reference(antigen = "SarsCoV2", epitope = "S2", org = "human", binding = TRUE) %>%
  filter(ref_publication != "LopezDeAssis_2023")

query <- PC_s2p_neg %>%
  filter(heavyCDR3_dist < cutoff) %>%
  column_to_rownames("barcodes") %>%
  .[,1:3]

find_publicBCR(
  query, 
  reference, 
  dist_method = "levenshtein",
  #heavyV = "heavyV",
  heavyCDR3 = "heavyCDR3", 
  output_dir = "./benchmark/test4/")

```